---
title: "Project 5"
description: |
  Analyses of Traffic Stops Across the USA
author: Laura Furtado Fernandes
date: April 29, 2025
format:
  html:
    code-fold: true
    code-summary: "Show the code"
execute:
  error: false  
  warning: false
---

```{r}

library(tidyverse)
library(RMariaDB)
library(DBI)
```

```{r}
con_traffic <- DBI::dbConnect(
  RMariaDB::MariaDB(),
  dbname = "traffic",
  host = Sys.getenv("TRAFFIC_HOST"),
  user = Sys.getenv("TRAFFIC_USER"),
  password = Sys.getenv("TRAFFIC_PWD")
)
```

The [Stanford Open Policing Project](https://openpolicing.stanford.edu/) compiled data from police stops from 42 different states across the United States. In this project, I analyze the racial make up of police stops in three major California cities—Los Angeles, San Diego, and San Francisco—to explore differences in stop patterns across locations and stop types. I will use SQL to access this data and wrangle it to examine my variables of interest: the race of the person stopped by police and the type of stop. I will limit my analyses to the year 2015.

```{sql}
#| connection: con_traffic
SELECT * FROM (
  SELECT subject_race, type,   
   SUM(1) AS num_stops,
   'los angeles' AS city
  FROM ca_los_angeles_2020_04_01
  WHERE type IS NOT NULL
    AND date BETWEEN '2015-01-01' AND '2015-12-31'
  GROUP BY  subject_race, type

  UNION ALL

  SELECT subject_race, type,   
    SUM(1) AS num_stops,
    'san diego' AS city
  FROM ca_san_diego_2020_04_01
  WHERE type IS NOT NULL
    AND date BETWEEN '2015-01-01' AND '2015-12-31'
  GROUP BY  subject_race, type

  UNION ALL

  SELECT subject_race, type,   
    SUM(1) AS num_stops,
   'san francisco' AS city
  FROM ca_san_francisco_2020_04_01
  WHERE type IS NOT NULL
    AND date BETWEEN '2015-01-01' AND '2015-12-31'
  GROUP BY  subject_race, type
) AS combined
ORDER BY num_stops DESC;
```

The table above displays the number and type of police stops by race across three California cities, ordered in descending fashion by total number of stops. The plot below visualizes these results. Los Angeles recorded the highest number of stops overall, with Hispanic individuals being stopped most frequently in both vehicular and pedestrian contexts. San Diego is a distant second in terms of overall stops, with San Francisco coming last in this comparison. In San Diego and San Francisco, white drivers are stopped most often, however, these results must be taken in consideration of the racial make up of the three cities.

According to the U.S. Census Bureau, Los Angeles has a Hispanic majority, with approximately 47% of its population identifying as Hispanic. In contrast, San Diego’s population is majority White, with 50% identifying as White. Similarly, in San Francisco, White individuals also make up the largest racial group, comprising about 50% of the population.

San Francisco and San Diego lack data on pedestrian stops which limits this comparison.

```{r}

df <- dbGetQuery(con_traffic, "
  SELECT subject_race, type,   
  SUM(1) AS num_stops,
  'los angeles' AS city
FROM ca_los_angeles_2020_04_01
WHERE type IS NOT NULL
  AND date BETWEEN '2015-01-01' AND '2015-12-31'
GROUP BY  subject_race, type

UNION ALL

SELECT subject_race, type,   
  SUM(1) AS num_stops,
  'san diego' AS city
FROM ca_san_diego_2020_04_01
WHERE type IS NOT NULL
  AND date BETWEEN '2015-01-01' AND '2015-12-31'
GROUP BY  subject_race, type

UNION ALL

SELECT subject_race, type,   
  SUM(1) AS num_stops,
  'san francisco' AS city
FROM ca_san_francisco_2020_04_01
WHERE type IS NOT NULL
  AND date BETWEEN '2015-01-01' AND '2015-12-31'
GROUP BY  subject_race, type
")

ggplot(df, aes(x = subject_race, y = num_stops, fill = type)) + 
  geom_col(position = "dodge") + 
  labs(
    title = "Number of Stops by Race and Stop Type in Los Angeles, \nSan Diego and San Francisco in 2015",
    x = "Subject Race",
    y = "Number of Stops",
    fill = "Stop Type"
  ) +
  theme_minimal() +
  facet_wrap( . ~ city) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```

Comparing within California is good to get a sense of within state differences in traffic stop patterns, however, these comparisons are not without limitations. For instance, Los Angeles is much more car reliant than San Francisco and also much more populous. San Diego might be more car-centric than San Francisco but Los Angeles experiences much higher traffic and has more extensive freeways than San Diego. Therefore, I decided to compare Los Angeles, California, to another highly car-centered city: Houston, Texas.

```{sql}
#| connection: con_traffic
SELECT * FROM (
  SELECT subject_race, type,   
    SUM(1) AS num_stops,
    'los angeles' AS city
  FROM ca_los_angeles_2020_04_01
  WHERE type IS NOT NULL
    AND date BETWEEN '2015-01-01' AND '2015-12-31'
  GROUP BY  subject_race, type

  UNION ALL

  SELECT subject_race, type,   
    SUM(1) AS num_stops,
   'houston' AS city
  FROM tx_houston_2023_01_26
  WHERE type IS NOT NULL
    AND date BETWEEN '2015-01-01' AND '2015-12-31'
  GROUP BY  subject_race, type
) AS combined
ORDER BY num_stops DESC;





```

The table above displays the number and type of police stops by race in Los Angeles, CA, and Houston, TX, ordered in descending order by total stops. The plot below visualizes this data. Compared to other California cities, Houston and Los Angeles show more similar overall stop volumes. In Houston, White individuals are the most frequently stopped demographic, followed by Black individuals. In contrast, Los Angeles shows comparable numbers of vehicular stops for Black and White individuals, but White individuals are stopped less often in pedestrian encounters.

Houston is missing data on pedestrian stops and on stops of Hispanic individuals which makes again limits comparisons.

```{r}
df <- dbGetQuery(con_traffic, "
  SELECT subject_race, type,   
    SUM(1) AS num_stops,
    'los angeles' AS city
  FROM ca_los_angeles_2020_04_01
  WHERE type IS NOT NULL
    AND date BETWEEN '2015-01-01' AND '2015-12-31'
  GROUP BY  subject_race, type

  UNION ALL

  SELECT subject_race, type,   
    SUM(1) AS num_stops,
   'houston' AS city
  FROM tx_houston_2023_01_26
  WHERE type IS NOT NULL
    AND date BETWEEN '2015-01-01' AND '2015-12-31'
  GROUP BY  subject_race, type
")

ggplot(df, aes(x = subject_race, y = num_stops, fill = type)) + 
  geom_col(position = "dodge") + 
  labs(
    title = "Number of Stops by Race and Stop Type in Los Angeles and Houston in 2015",
    x = "Subject Race",
    y = "Number of Stops",
    fill = "Stop Type"
  ) +
  theme_minimal() +
  facet_wrap( . ~ city) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

In this project, I used SQL to analyze police stop data from the Stanford Open Policing Project, focusing on racial patterns and stop types in three major California cities—Los Angeles, San Diego, and San Francisco—as well as a cross-state comparison with Houston, Texas. Within California, Los Angeles showed the highest number of stops, but the distribution of stops by race varied between cities. While comparing Los Angeles and Houston, I ran into an issue of missing data on the ethnicity (whether person stopped was of Hispanic origin) of the person stopped which limited comparisons with Los Angeles, which has a large hispanic population. I did see that Houston and Los Angeles have similar traffic stop counts, with makes sense given the car-centered culture of both cities.
