---
title: "Hollywood Age Gaps"
description: |
  Analysis of Hollywood Age Gaps Over Time
author: Laura Furtado Fernandes
date: February 16, 2025
format: html
---

```{r, include=FALSE}
library(readr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
age_gaps <- read_csv("tidytuesday/age_gaps.csv")

age_gaps$same_sex <- ifelse(age_gaps$character_1_gender == age_gaps$character_2_gender, "yes", "no")


```

```{r, echo = FALSE}
library(RColorBrewer)

#| fig-alt: "Scatter plot of age gaps between Hollywood couples over time coloured by whether the couple was of the same gender or not, where there are many more non-same gender couples than same gender couples."
#| label: age-gap-plot
#| fig-cap: "Hollywood Age Gaps Over Time"


ggplot(age_gaps, aes(x = release_year, y = age_difference)) + 
  geom_point(aes(colour = same_sex), alpha = 0.7) + 
  scale_colour_brewer(palette="Dark2") + 
  labs(title = "Age Differences in Hollywood Movie Couples Over the Years", 
       subtitle = "From 1931 to 2025", 
       colour = "Is it a same \ngender couple?", 
       x = "", 
       y = "") + 
  theme_minimal()


```

**This plot shows that couple age gaps have remained pretty consistent over time and that the data was skewed towards more recent movies. It also shows that there are many more non-same gender couples than same gender couples in Hollywood movies but that the same gender couples are equally spread out along the age gap axis. Same gender couples only appeared in this Hollywood data set after the year 2000.**

Citations:

This data comes from [Hollywood Age Gap](https://hollywoodagegap.com/) via [Data Is Plural](https://www.data-is-plural.com/archive/2018-02-07-edition/).

Data Science Learning Community (2024). Tidy Tuesday: A weekly social data project. [https://tidytues.day](https://tidytues.day/)

<https://github.com/rfordatascience/tidytuesday/tree/main/data/2023/2023-02-14>
